---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

### __partiesAL__:  Legislative Parties of the Americas 1925 - 2019


*Aníbal Pérez-Liñán, Nicolás Schmidt, Daniela Vairo*



<!-- badges: start -->
`r badger::badge_devel("Nicolas-Schmidt/partiesAL", "orange")`
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)

<!-- badges: end -->



## Description
This package offers two interconnected datasets. The first dataset, named __`parties`__, focuses on political parties as the unit of analysis. It covers a span of 95 years (1925-2019) and includes 21 countries from the Americas. This group comprises 20 Latin American countries and the United States. The dataset provides comprehensive information about political parties in each year of the congress for these countries.

The second dataset, referred to as __`parties2`__, serves as an aggregate of the aforementioned data, consolidating it into a country-year unit. In addition to the party-specific data, this second database includes a diverse range of systemic variables. These variables offer valuable insights into the political landscape, such as the effective number of parties in each chamber of the congress.

There are two methods available for accessing the data. The first option involves installing the R package, `partiesAL`, which can be done by following the step-by-step instructions provided below. This method is suitable for users who prefer a conventional installation from a repository.

Alternatively, for users who prefer accessing the data outside of R, there are three available choices: csv, xlsx, and dta formats. These formats allow users to work with the data in their preferred software or programming environment.

To gain a comprehensive understanding of the variables and their meanings within the databases, the codebook can be accessed through the following link. The codebook provides detailed explanations and descriptions of the variables used in the datasets.


##### Data download

* __`parties`__: [csv](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures/parties/export/csv)  [xlsx](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures/parties/export/xlsx)  [dta](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures/parties/export/dta) 
* __`parties2`__: [csv](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures/parties2/export/csv)  [xlsx](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures/parties2/export/xlsx)  [dta](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures/parties2/export/dta) 
* __`codebook`__: [here](http://www.politicaygobierno.cide.edu/index.php/pyg/article/view/1666/1101)
* __`paper`__: [here](http://www.politicaygobierno.cide.edu/index.php/pyg/article/view/1666/1101) and [here](https://www.tandfonline.com/doi/abs/10.1080/13510347.2019.1566321)


## `parties` and `parties2`

The structure of the databases is as follows

```{r}
str(partiesAL::parties)
```


```{r}
str(partiesAL::parties2)
```




## Installation

To install the development version from GitHub:

``` r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("Nicolas-Schmidt/partiesAL")
```


## Example
```{r, echo=F, message=FALSE, warning=FALSE}
library(tidyverse)

```

In the following example, the `parties` database is used to visualize the temporal evolution of the number of political parties in the congresses of each country.



```{r, fig.width = 10, fig.height = 12}
library(partiesAL)

nparty <- 
    partiesAL::parties %>% 
    select(ccode, pty_code, year, legis) %>% 
    filter(legis %in% c(1,3)) %>% 
    distinct() %>%
    select(-pty_code, - legis) %>% 
    split(., .$ccode) %>% 
    lapply(., table) %>% 
    lapply(., as.data.frame) %>% 
    do.call('rbind', .) 


ggplot(nparty, aes(x = as.numeric(as.character(year)), y = Freq))+
    geom_bar(stat="identity", width = NULL, fill = "#279F00", color = "black") + 
    facet_wrap(~ccode, ncol = 3) + 
    theme_minimal() +
    labs(x = "", y = "")


```
