---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

### __partiesAL__:  Legislative Parties of the Americas 1925 - 2019


*Aníbal Pérez-Liñán, Nicolás Schmidt, Daniela Vairo*



<!-- badges: start -->
`r badger::badge_devel("Nicolas-Schmidt/partiesAL", "orange")`
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)

<!-- badges: end -->



## Description

This package provides two related datasets. The first one, called __`parties`__, is a dataset where the unit of analysis is the political party in each year in the congress for 21 countries of the Americas (20 Latin American countries plus the United States) in the period from 1925 to 2019. The second dataset (__`parties2`__) is an aggregation of this data, compiled into a unit that becomes country-year. This second database provides an additional set of systemic variables such as the effective number of parties in each chamber of the congress.


The countries included in the database are:
```{r}
unique(partiesAL::parties$ccode)
```


* There are two ways to access the data. One option is to install this R package (`partiesAL`) by following the detailed instructions below for a conventional installation from a repository. The second option, for users who prefer to access the data outside of R, provides three available choices: [__csv__](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures), [__xlsx__](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures),  [__dta__](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures).              


* The codebook for the databases can be accessed through this [__link__](https://github.com/Nicolas-Schmidt/partiesAL/tree/master/man/figures).




## `parties` and `parties2`

The structure of the databases is as follows

```{r}
str(partiesAL::parties)
```


```{r}
str(partiesAL::parties2)
```




## Installation

To install the development version from GitHub:

``` r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("Nicolas-Schmidt/partiesAL")
```


## Example
```{r, echo=F, message=FALSE, warning=FALSE}
library(tidyverse)

```

In the following example, the `parties` database is used to visualize the temporal evolution of the number of political parties in the congresses of each country.



```{r, fig.width = 10, fig.height = 12}
library(partiesAL)

nparty <- 
    partiesAL::parties %>% 
    select(ccode, pty_code, year, legis) %>% 
    filter(legis %in% c(1,3)) %>% 
    distinct() %>%
    select(-pty_code, - legis) %>% 
    split(., .$ccode) %>% 
    lapply(., table) %>% 
    lapply(., as.data.frame) %>% 
    do.call('rbind', .) 


ggplot(nparty, aes(x = as.numeric(as.character(year)), y = Freq))+
    geom_bar(stat="identity", width = NULL, fill = "red", color = "black")+ 
    facet_wrap(~ccode, ncol = 3) + 
    theme_minimal() +
    labs(x = "", y = "")


```
