---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# partiesAL

*Aníbal Pérez-Liñán, Nicolás Schmidt, Daniela Vairo*
<!-- badges: start -->
`r badger::badge_devel("Nicolas-Schmidt/partiesAL", "orange")`
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)

<!-- badges: end -->


## Description

This package provides two related datasets. The first one, called `parties`, is a dataset where the unit of analysis is the political party in each year in the congress of 21 Latin American countries from 1925 to 2020. The second dataset (`parties2`) is an aggregation of these data, compiled into a unit that becomes country-year

The countries included in the database are:
```{r}
unique(partiesAL::parties$ccode)
```


* There are two ways to access the data. One option is to install this R package according to the instructions provided below for a conventional installation from a repository. The second option, for users who prefer accessing the data in xlsx format, can find the data at this link.


* The codebook for the databases can be accessed through this link.




## `parties` and `parties2`

The structure of the databases is as follows.

```{r}
str(partiesAL::parties)
```


```{r}
str(partiesAL::parties2)
```




## Installation

To install the development version from GitHub:

``` r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("Nicolas-Schmidt/partiesAL")
```


## Example
```{r, echo=F, message=FALSE, warning=FALSE}
library(tidyverse)

```

In the following example, the `parties` database is used to visualize the temporal evolution of the number of political parties in the congresses of each country.



```{r, fig.width = 10, fig.height = 12}
library(partiesAL)

nparty <- 
    partiesAL::parties %>% 
    select(ccode, pty_code, year, legis) %>% 
    filter(legis %in% c(1,3)) %>% 
    distinct() %>%
    select(-pty_code, - legis) %>% 
    split(., .$ccode) %>% 
    lapply(., table) %>% 
    lapply(., as.data.frame) %>% 
    do.call('rbind', .) 


ggplot(nparty, aes(x = as.numeric(as.character(year)), y = Freq))+
    geom_bar(stat="identity", width = NULL, fill = "red", color = "black")+ 
    facet_wrap(~ccode, ncol = 3) + 
    theme_minimal() +
    labs(x = "", y = "")


```
